---
title: "Primera Práctica"
author: "Equipo 8: -Flavio Cesar Palacios Salas - Berlin Isabel Preciado Escalera - Daniel Orozco Gomez"
date: "07-02-2022"
output: rmdformats::readthedown
---

# Simulación de procesos

## Proceso de Wiener

En este código simularemos un proceso de Wiener Generalizado y observaremos la variable que siga un comportamiento de ese tipo

$$
dW = \epsilon\sqrt{dt}
$$

**Donde:**

$\epsilon ~ N(0,1)$

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
### Carga de paqueterías para cargar datos y graficar
library(tidyverse)
```

```{r echo=FALSE, include=FALSE}
set.seed(123)
sim <- 100000
dt <- 1/sim
epsilon <- rnorm(n=sim, mean = 0, sd = 1)
dw <- epsilon*sqrt(dt)
```

```{r}
datos <- tibble(t=seq(1/sim,1,by=dt),dw)
ggplot(datos)+
  geom_line(aes(x=t,y=dw))
```

```{r echo=FALSE}
datos <- tibble(t = seq(dt,1,by= dt),dw)
ggplot(datos)+
  geom_line(aes(x=t,y=dw), color = "#C70039")+
  labs(x= "Tiempo",y = "dW", title = "Proceso de Wiener")
```

## Proceso de Wiener Generalizado

¿cómo sería W? $$W_t = W_{t-1} +dW_t$$

```{r}
W <- 0
W[1] <- 0
for (t in 2:sim){
  W[t] = W[t-1]+dw[t]
}
datos <- mutate(datos,W=W)
```

```{r, echo=FALSE, message=FALSE}
ggplot(datos)+
  geom_line(aes(x=t, y=W), color = "#C70039")+
  geom_smooth(aes(x=t,y=W),method="lm", color = "black")
```

## Proceso de Wiener Generalizado
En este caso
$$ dX = adt+bdw$$

```{r}
a <- 0.3
b <- 0.15
dx <-  a*dt+b*dw
plot(dx,type = 'l')
```
Ahora X, sería
$$X_t = X_{t-1}+dX_t$$
```{r}
X <- 0
X[1] <-0
for (t in 2:sim){
  X[t] = X[t-1]+dx[t]
}
datos <- mutate(datos,X=X)
```

```{r,echo = FALSE, message=FALSE}
ggplot(datos)+
  geom_line(aes(x=t, y=X), color = "#C70039")+
  geom_smooth(aes(x=t,y=X),method="lm", color = "Black")
```

## Resolviendo un problema
Una determinada variable sigue un proceso de WG con una drift rate de 2 y una tasa de varianza de 4. Ambas con unidad de tiempo semestral. Si el día de hoy la variable X toma el valor de 50.
1 año
$$dX = 4+2\sqrt{2} \epsilon$$
$$X=50+dX$$
$$X=54+2\sqrt{2}\epsilon$$
¿Cuál es el valor esperado y la varianza de X dentro de 1 año?
```{r}
dxa <- 4+2*sqrt(2)*epsilon
Xa <- 50+dxa
```

```{r}
mean(Xa)
var(Xa)
```

¿Cuáles son las probabilidades de que el cambio en X sea mayor a cero?
```{r}
mean(dxa<0)
```

¿Cuáles son las probabilidades de que el cambio en X sea mayor a cero?
```{r}
mean((50<Xa)&(Xa<=51))
```

## Problemas

### Primero
I) A company’s cash position, measured in millions of dollars, follows a generalized. Wiener process with a drift rate of 0.5 per quarter and a variance rate of 4.0 per quarter. How high does the company’s initial cash position have to be for the company to have a less than 5% chance of a negative cash position by the end of 1 year? To solve this, follow the next steps;

1) Identify a and b for the generalized Wiener Process.

Para el proceso de Wiener Generalizado se tiene la siguiente forma

$$ dX = adt+bdw$$

dado que a corresponde al drift rate y b la razón de varianza entonces a corresponde a 0.5 y b a 4.0

2) Write the equation dx using a and b. That equation helps us to solve the problem and is specific for this problem.

$$ dX = 0.5dt+4dw$$

3) Using 2) write the equation dx when dt=4 (fourth quarters or one year)

$$ dX = 0.5(4)+4\sqrt(4)\epsilon$$
4) Suppose that the initial cash position is Xo. Write the equation for X (not for dx)
$$ X = X_0 +0.5(4)+4\sqrt(4)\epsilon$$

5) Find the expected value and variance of X
$$ E[X] = E[X_0 +0.5(4)+4\sqrt(4)\epsilon]$$
$$ E[X] = E[X_0] +E[0.5(4)]+E[4\sqrt(4)\epsilon]$$
$$ E[X] = X_0 +0.5(4)+0$$
$$ E[X] = X_0 +2$$

$$ Var[X] = Var[X_0 +0.5(4)+4\sqrt(4)\epsilon]$$
$$ Var[X] = Var[X_0] Var[+0.5(4)]+Var[4\sqrt(4)\epsilon]$$
$$ Var[X] = 0+0+Var[4\sqrt(4)\epsilon]$$
$$ Var[X] = 64Var[\epsilon]$$
$$ Var[X] = 64$$
6) Using 5) find P(X<0). Normalize and use normal standard distribution. If you are doing good, you should get a result that depends on Xo. Now your job is to find this value. 

$$P(X<0)=\frac{0-X_0-2}{2\sqrt{2}}$$

7) Equalize de result on 6) to -1.65 (because you want less than 5% chance of a negative position). And solve for Xo.
$$-1.65=\frac{-X_0-2}{2\sqrt{2}}$$
$$-1.65(2\sqrt{2})=-X_0-2$$
$$-1.65(2\sqrt{2})+2=-X_0$$
$$X_0=1.65(2\sqrt{2})-2$$

$$X_0 = 2.6669$$
8) Use R to prove your answer. Simulate the equation X=Xo+adt+bdW using the Xo you found on 7) with dt=4 and a,b for the problem. Once you have at least 10000 simulations for X, calculate its mean and variance. Finally, use pnorm(0, mean(X), sd(X)) to find the probability that X (initial cash position) has to be less than zero. 

```{r}
epsilon <- rnorm(n=10000, mean = 0, sd = 1)
X <- 2.6669+2+2*sqrt(2)*epsilon
```

```{r}
mean(X)
var(X)
```

¿Cuáles son las probabilidades de que el cambio en X sea mayor a cero?
```{r}
mean(X<0)
```
Se puede observar que el valor es muy cercano al 5% de probabilidad del problema, no se acerca estrictamente por el número de simulaciones, si se aumentaran este número sería más cercano, sin embargo el valor se acerca

### Segundo
II) A company’s cash position, measured in millions of dollars, follows a generalized Wiener process with a drift rate of 0.1 per month and a variance rate of 0.16 per month. The initial cash position is 2.0.

(a) What are the probability distributions of the cash position after 1 month, 6 months, and 1 year?

Considerando los datos anteriores el anterior poblema se puede modelar de la siguiente forma:
$$ X = 2 +0.1dt+0.16dW$$
Para determinar las distirbuciones de probabilidad para cada uno de los periodos hay que calcular las esperanzas y varianzas para cada uno.

Para dt = 1 mes
$$ X_1 = 2 +0.1(1)+0.16\epsilon$$
Haciendo los calculos
$$E[X_1]=2.1$$
$$Var[X_1]=(0.16)^2$$
Por tanto para el primer mes sigue una distribución $N(2.1,0.16)$

Para dt = 6 meses
$$ X = 2 +0.1(6)+0.16\sqrt{6}\epsilon$$
$$E[X_6]=2.6$$
$$Var[X_6]=(0.3919)^2$$
Por tanto para el primer mes sigue una distribución $N(2.6,0.3919)$

Para dt = 12 meses
$$ X = 2 +0.1(12)+0.16\sqrt{12}\epsilon$$
$$E[X_{12}]=3.2$$
$$Var[X_{12}]=(0.5543)^2$$
Por tanto para el primer mes sigue una distribución $N(3.2,0.5543)$

(b) What are the probabilities of a negative cash position at the end of 6 months and 1 year?

$$P(X<0)=P(\frac{0-2.6}{0.16})=P(\frac{-2.6}{0.16})=P(-6.63)=1.6784(10)^{-11}$$

$$P(X<0)=P(\frac{0-2.6}{0.16})=P(\frac{-3.2}{0.5543})=P(-5.773)=3.8936(10)^{-9}$$
(c) Use R to simulate the process and make a graph where you can see the relation between t form 1 to 40 months and the evolution of the probability 
Hint: to solve this problem, use the same steps on I)
```{r}
t = seq(from = 1, to = 40, by = 1)
X_mean = 2+0.1*t
X_68u = 2 +0.1*t+0.16*t**0.5
X_68l = 2 +0.1*t-0.16*t**0.5
X_95u = 2 +0.1*t+2*0.16*t**0.5
X_95l = 2 +0.1*t-2*0.16*t**0.5
X_99u = 2 +0.1*t+3*0.16*t**0.5
X_99l = 2 +0.1*t-3*0.16*t**0.5
```

```{r}
ggplot()+
  geom_line(aes(x=t,y=X_68u,color = "68%"))+
  geom_line(aes(x=t,y=X_68l,color= "68%"))+
  geom_line(aes(x=t,y=X_95u, color = "95%"))+
  geom_line(aes(x=t,y=X_95l, color = "95%"))+
  geom_line(aes(x=t,y=X_99u, color = "99%"))+
  geom_line(aes(x=t,y=X_99l, color = "99%"))+
  geom_line(aes(x=t,y=X_mean))+
  labs(x= "Mes(t)",y = "X", title = "Evolución de la distribución de la posición segun intervarlos de confianza")+
  theme_classic()
```





