---
title: "Primera Práctica"
author: "Equipo 8: -Flavio Cesar Palacios Salas - Berlin Isabel Preciado Escalera - Daniel Orozco Gomez"
date: "07-02-2022"
output: rmdformats::readthedown
---

# Simulación de procesos

## Proceso de Wiener

En este código simularemos un proceso de Wiener Generalizado y observaremos la variable que siga un comportamiento de ese tipo

$$
dW = \epsilon\sqrt{dt}
$$

**Donde:**

$\epsilon ~ N(0,1)$

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
### Carga de paqueterías para cargar datos y graficar
library(tidyverse)
```

```{r echo=FALSE, include=FALSE}
set.seed(123)
sim <- 100000
dt <- 1/sim
epsilon <- rnorm(n=sim, mean = 0, sd = 1)
dw <- epsilon*sqrt(dt)
```

```{r}
datos <- tibble(t=seq(1/sim,1,by=dt),dw)
ggplot(datos)+
  geom_line(aes(x=t,y=dw))
```

```{r echo=FALSE}
datos <- tibble(t = seq(dt,1,by= dt),dw)
ggplot(datos)+
  geom_line(aes(x=t,y=dw), color = "#C70039")+
  labs(x= "Tiempo",y = "dW", title = "Proceso de Wiener")
```

## Proceso de Wiener Generalizad

¿cómo sería W? $$W_t = W_{t-1} +dW_t$$

```{r}
W <- 0
W[1] <- 0
for (t in 2:sim){
  W[t] = W[t-1]+dw[t]
}
datos <- mutate(datos,W=W)
```

```{r, echo=FALSE, message=FALSE}
ggplot(datos)+
  geom_line(aes(x=t, y=W), color = "#C70039")+
  geom_smooth(aes(x=t,y=W),method="lm", color = "black")
```

## Proceso de Wiener Generalizado
En este caso
$$ dX = adt+bdw$$

```{r}
a <- 0.3
b <- 0.15
dx <-  a*dt+b*dw
plot(dx,type = 'l')
```
Ahora X, sería
$$X_t = X_{t-1}+dX_t$$
```{r}
X <- 0
X[1] <-0
for (t in 2:sim){
  X[t] = X[t-1]+dx[t]
}
datos <- mutate(datos,X=X)
```

```{r,echo = FALSE, message=FALSE}
ggplot(datos)+
  geom_line(aes(x=t, y=X), color = "#C70039")+
  geom_smooth(aes(x=t,y=X),method="lm", color = "Black")
```

## Resolviendo un problema
Una determinada variable sigue un proceso de WG con una drift rate de 2 y una tasa de varianza de 4. Ambas con unidad de tiempo semestral. Si el día de hoy la variable X toma el valor de 50.
1 año
$$dX = 4+2\sqrt{2} \epsilon$$
$$X=50+dX$$
$$X=54+2\sqrt{2}\epsilon$$
¿Cuál es el valor esperado y la varianza de X dentro de 1 año?
```{r}
dxa <- 4+2*sqrt(2)*epsilon
Xa <- 50+dxa
```

```{r}
mean(Xa)
var(Xa)
```

¿Cuáles son las probabilidades de que el cambio en X sea mayor a cero?
```{r}
mean(dxa<0)
```

¿Cuáles son las probabilidades de que el cambio en X sea mayor a cero?
```{r}
mean((50<Xa)&(Xa<=51))
```

## Problemas

### Primero
I) A company’s cash position, measured in millions of dollars, follows a generalized. Wiener process with a drift rate of 0.5 per quarter and a variance rate of 4.0 per quarter. How high does the company’s initial cash position have to be for the company to have a less than 5% chance of a negative cash position by the end of 1 year? To solve this, follow the next steps;

1) Identify a and b for the generalized Wiener Process.
2) Write the equation dx using a and b. That equation helps us to solve the problem and is specific for this problem. 
3) Using 2) write the equation dx when dt=4 (fourth quarters or one year)
4) Suppose that the initial cash position is Xo. Write the equation for X (not for dx)
5) Find the expected value and variance of X
6) Using 5) find P(X<0). Normalize and use normal standard distribution. If you are doing good, you should get a result that depends on Xo. Now your job is to find this value. 
7) Equalize de result on 6) to -1.65 (because you want less than 5% chance of a negative position). And solve for Xo.
8) Use R to prove your answer. Simulate the equation X=Xo+adt+bdW using the Xo you found on 7) with dt=4 and a,b for the problem. Once you have at least 10000 simulations for X, calculate its mean and variance. Finally, use pnorm(0, mean(X), sd(X)) to find the probability that X (initial cash position) has to be less than zero. 

### Segundo
II) A company’s cash position, measured in millions of dollars, follows a generalized Wiener process with a drift rate of 0.1 per month and a variance rate of 0.16 per month. The initial cash position is 2.0.

(a) What are the probability distributions of the cash position after 1 month, 6 months, and 1 year?
(b) What are the probabilities of a negative cash position at the end of 6 months and 1 year?
(c) Use R to simulate the process and make a graph where you can see the relation between t form 1 to 40 months and the evolution of the probability 
Hint: to solve this problem, use the same steps on I)